---
title: "Relating module to trait"
author: "koji shiota"
date: "2020/7/11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Script")
```

```{r}
library(tidyverse)
library(WGCNA)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
```



```{r}
lnames = load("../Output/001_NormalizeRawData/SLE_GSE138458-01-dataInput.RData")
print(lnames)
```


```{r}
datTraits <- datTraits %>% 
  # filter(description %in% rownames(datExpr)) %>% 
  dplyr::select(-c(ID_REF, case_control, sledai_group))

gene_annotation <- read_tsv("../Data/GPL10558-50081.txt", comment = "#", col_names = TRUE)
```


```{r}
constructedNetwork_paths <- list.files(path = "../Output/002_NetworkConstruction", pattern = "SLE_GSE138458-02-networkConstruction_.*.RData", full.names = TRUE)
constructedNetwork_paths#[[1]]
```


```{r}
constructedNetwork = "../Output/002_NetworkConstruction/SLE_GSE138458-02-networkConstruction_unsigned_4.RData"
#for (constructedNetwork in constructedNetwork_paths){
condition = basename(constructedNetwork) %>% str_extract(pattern = "(un)?signed_[0-9](_merged)?")
print(condition)
output_dir = paste0("../Output/003_RelatingModuleToTrait/", condition, "/")
if (!dir.exists(output_dir)){
  dir.create(output_dir)}
lnames = load(file = constructedNetwork)
print(lnames)
```


```{r}
## 3.a Quantifying module–trait associations
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
```



#####################################################

reference
# III. Using simulated data to evaluate different module detection methods and gene screening approaches

## 6. Relating modules and module eigengenes to external data

### 6.a Representing modules by eigengenes and relating eigengenes to one another
### To get a sense of how related the modules are one can summarize each module by its eigengene (first principal component).
```{r}
signif(cor(MEs, use="p"), 2)
```

```{r}
dissimME=(1-t(cor(MEs, method="p")))/2
hclustdatME=hclust(as.dist(dissimME), method="average" )
# Plot the eigengene dendrogram
par(mfrow=c(1,1))
plot(hclustdatME, main="Clustering tree based of the module eigengenes")
```


## 6.a.1 Pairwise scatter plots of the samples (arrays) along the module eigengenes
モジュールが多すぎるのでplotMEpairsではよく見えない
```{r}
#sizeGrWindow(8,9)
plotMEpairs(MEs,y=datTraits$case_control_levels)
```

## 6.c Diagnostics: heatmap plots of module expression
```{r}
sizeGrWindow(8,9)
par(mfrow=c(3,1), mar=c(1, 2, 4, 1))
which.module="green"
plotMat(t(scale(datExpr[,moduleColors==which.module ]) ),
        nrgcols=30,rlabels=T,
        clabels=T,rcols=which.module,
        title=which.module )

which.module="red"
plotMat(t(scale(datExpr[,moduleColors==which.module ]) ),
        nrgcols=30,rlabels=T,
        clabels=T,rcols=which.module,
        title=which.module )
```


## 6.d Diagnostics: displaying module heatmap and the eigengene
```{r}
sizeGrWindow(8,7);
which.module="red"
ME=MEs[, paste("ME",which.module, sep="")]
par(mfrow=c(2,1), mar=c(0.3, 5.5, 3, 2))
plotMat(t(scale(datExpr[,moduleColors==which.module ]) ),
        nrgcols=30,rlabels=F,rcols=which.module,
        main=which.module, cex.main=2)
par(mar=c(5, 4.2, 0, 0.7))
barplot(ME, col=which.module, main="", cex.main=2,
ylab="eigengene expression",xlab="array sample")
```








#####################################################


```{r}
moduleTraitCor = cor(MEs, datTraits, use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
```


```{r}
sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
#png(file = paste0(output_dir,"Module-trait_relationships.png"), width = 720, height = 480)
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(datTraits),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
#dev.off()

```

```{r}
GS1=as.numeric(cor(datTraits$sledai_group_levels,datExpr, use="p"))
GeneSignificance=abs(GS1)
# Next module significance is defined as average gene significance.
ModuleSignificance=tapply(GeneSignificance, moduleColors, mean, na.rm=T)
ModuleSignificance
```

```{r}
sizeGrWindow(8,7)
par(mfrow = c(1,1))
plotModuleSignificance(GeneSignificance,moduleColors)
```











# ここからはパスウェイ解析とかになるから上までのModuleとTraitの相関解析とは毛色が若干異なる
# → 次のスクリプトに回したほうが良さげ？
```{r}
pvalue_threshold = 0.05

sum(moduleTraitPvalue[,"case_control_levels"] < pvalue_threshold)
significant_module <- moduleTraitPvalue[moduleTraitPvalue[,"case_control_levels"] < pvalue_threshold,] %>% 
  rownames() %>% 
  str_sub(start = 3)

gene_info <- tibble(
  probeID = colnames(datExpr),
  module = moduleColors
) %>% 
  left_join(gene_annotation %>% dplyr::select(c(ID, Entrez_Gene_ID, Symbol, Synonyms)), by = c("probeID" = "ID"))


# Enrichment 

groupby_enrichment <- gene_info %>% 
  group_by(module) %>% 
  nest() %>% 
  filter(module != "grey", module %in% significant_module) %>% 
  #head(2) %>% 
  mutate(GOenrich = map(data, .f = function(data){enrichGO(data$Entrez_Gene_ID, OrgDb = org.Hs.eg.db, ont = "ALL", readable = TRUE)})#,
         #GOenrich_readable = map(data, .f = function(data){enrichGO(data$Entrez_Gene_ID, OrgDb = org.Hs.eg.db, ont = "ALL", readable = TRUE)})
  )

map2(groupby_enrichment$GOenrich , groupby_enrichment$module,
     .f = function(GOenrich, module){
       bar = barplot(GOenrich)
       ggsave(filename = paste0(output_dir, module,"_barplot",".pdf"), plot = bar,
              width = 300,height = 200,units = "mm")
       GOenrich@result %>% 
         write_csv(paste0(output_dir, module,"_result",".csv"))
     }
)
#}
```









```{r}
constructedNetwork_paths <- list.files(path = "../Output/002_NetworkConstruction",
                                       pattern = "SLE_GSE138458-02-networkConstruction_.*.RData", full.names = TRUE)

for (constructedNetwork in constructedNetwork_paths){
  condition = basename(constructedNetwork) %>% str_extract(pattern = "(un)?signed_[0-9](_merged)?")
  print(condition)
  output_dir = paste0("../Output/003_RelatingModuleToTrait//", condition, "/")
  if (!dir.exists(output_dir)){
    dir.create(output_dir)}
  lnames = load(file = constructedNetwork)
  print(lnames)
  
  
  
  ## 3.a Quantifying module–trait associations
  # Define numbers of genes and samples
  nGenes = ncol(datExpr);
  nSamples = nrow(datExpr);
  # Recalculate MEs with color labels
  MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
  MEs = orderMEs(MEs0)
  moduleTraitCor = cor(MEs, datTraits, use = "p");
  moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
  
  sizeGrWindow(10,6)
  # Will display correlations and their p-values
  textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                     signif(moduleTraitPvalue, 1), ")", sep = "");
  dim(textMatrix) = dim(moduleTraitCor)
  par(mar = c(6, 8.5, 3, 3));
  # Display the correlation values within a heatmap plot
  png(file = paste0(output_dir,"Module-trait_relationships.png"), width = 720, height = 480)
  labeledHeatmap(Matrix = moduleTraitCor,
                 xLabels = names(datTraits),
                 yLabels = names(MEs),
                 ySymbols = names(MEs),
                 colorLabels = FALSE,
                 colors = blueWhiteRed(50),
                 textMatrix = textMatrix,
                 setStdMargins = FALSE,
                 cex.text = 0.5,
                 zlim = c(-1,1),
                 main = paste("Module-trait relationships"))
  dev.off()
  
  pvalue_threshold = 0.05
  
  sum(moduleTraitPvalue[,"case_control_levels"] < pvalue_threshold)
  significant_module <- moduleTraitPvalue[moduleTraitPvalue[,"case_control_levels"] < pvalue_threshold,] %>% 
    rownames() %>% 
    str_sub(start = 3)
  
  gene_info <- tibble(
    probeID = colnames(datExpr),
    module = moduleColors
  ) %>% 
    left_join(gene_annotation %>% dplyr::select(c(ID, Entrez_Gene_ID, Symbol, Synonyms)), by = c("probeID" = "ID"))
  
  
  # Enrichment 
  
  groupby_enrichment <- gene_info %>% 
    group_by(module) %>% 
    nest() %>% 
    filter(module != "grey", module %in% significant_module) %>% 
    #head(2) %>% 
    mutate(GOenrich = map(data, .f = function(data){enrichGO(data$Entrez_Gene_ID, OrgDb = org.Hs.eg.db, ont = "ALL", readable = TRUE)})#,
           #GOenrich_readable = map(data, .f = function(data){enrichGO(data$Entrez_Gene_ID, OrgDb = org.Hs.eg.db, ont = "ALL", readable = TRUE)})
    )
  
  map2(groupby_enrichment$GOenrich , groupby_enrichment$module,
       .f = function(GOenrich, module){
         bar = barplot(GOenrich)
         ggsave(filename = paste0(output_dir, module,"_barplot",".pdf"), plot = bar,
                width = 300,height = 200,units = "mm")
         GOenrich@result %>% 
           write_csv(paste0(output_dir, module,"_result",".csv"))
       }
  )
}
```



